---
title: "p8105_hw6_yw4199"
author: "Yaduo Wang"
date: "2023-11-29"
output: github_document
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
```

```{r setup, include=FALSE}
# general settings
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	fig.height = 6,
	fig.width = 8,
	message = FALSE,
	warning = FALSE,
	out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 2
```{r warning=FALSE}
# download the data
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


```{r warning=FALSE}
# use the bootstrap to examine the distribution under repeated sampling
boot_straps = weather_df|> 
  modelr::bootstrap(n = 5000, id = "ID") |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results_estimate = map(models, broom::tidy),
    extr_r = map(models, broom::glance)) |> 
  select(-strap, -models) 

# produce estimates of log of product of beta and r-squared
beta_calculated = boot_straps |> 
  unnest(results_estimate) |> 
  group_by(ID) |> 
  summarize(product = prod(estimate)) |> 
  mutate(log_product = log(product))

r_square = boot_straps |> 
  unnest(extr_r) |> 
  group_by(ID) |> 
  summarize(r_squared = mean(r.squared))
```

```{r}
# identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval 
log_product_beta_ci = beta_calculated |> 
  summarize(
    ci_lower = quantile(log_product, 0.025, na.rm = TRUE),
    ci_upper = quantile(log_product, 0.975, na.rm = TRUE)
  )
r_square_ci = r_square |> 
  summarize(
    ci_lower = quantile(r_squared, 0.025, na.rm = TRUE),
    ci_upper = quantile(r_squared, 0.975, na.rm = TRUE)
  )
```

```{r}
# Plot for log of product beta distribution
beta_calculated |>
  ggplot(aes(x = log_product)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  theme_minimal() +
  labs(x = "Log of Product Beta", y = "Frequency", 
       title = "Distribution of Log of Product Beta")

# Plot for r_squared distribution
r_square |> 
  ggplot(aes(x = r_squared)) +
  geom_histogram(fill = "lightblue",color = "black") +
  theme_minimal() +
  labs(x = "r_squared", y = "Frequency", 
       title = "Distribution of R_squared")
```

The distribution of the log of the product of the beta shows that most values are centered around -2.5 to -3.25. The distribution is left skewed and this feature might be related to the outliers in the bootstrap sample. The distribution of the r-squared is more normally distributed and the values are centered around 0.92. 

## Problem 3
```{r}
#Load and clean the data for regression analysis.
birthweight <- read.csv("DATA/birthweight.csv")
birthweight = 
  birthweight |> 
#convert numeric to factor
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  )
#check for missing data
colSums(is.na(birthweight))
```

After we loading the data, and convert some numeric variables to factor variables as necessary, we check whether there exist missing values. As shown in the results, there is no missing values in each variable. 

```{r}
#Propose a regression model for birth weight with all variables as predictor. 
fit_all = lm(bwt ~., data = birthweight)
summary(fit_all)
```

Select predictors that is significant with small p-value from the summary of the model shown above.

```{r}
# Fit more accurate model. 
fit_selected = lm(bwt ~ babysex + bhead + 
                    blength + delwt + gaweeks +
                    smoken, 
                  data = birthweight)
summary(fit_selected)
```

```{r}
# plot of model residuals against fitted values
birthweight |> 
  modelr::add_residuals(fit_selected) |> 
  modelr::add_predictions(fit_selected) |> 
  ggplot(aes(x = pred, y=resid)) + 
  geom_point(alpha = 0.5) + 
  geom_hline(yintercept = 0) +
  labs(x = "Fitted values", y = "Residuals", 
       title = "Model Residuals vs. Fitted Values")
```

```{r}
# cross validation between two models 
cv_df =crossv_mc(birthweight, 100)
cv_df = 
  cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df |> 
  mutate(
    mine_model = map(train, \(df) lm(bwt ~ babysex + bhead + 
                  blength + delwt + gaweeks + smoken, data = df)),
# One using length at birth and gestational age as predictors (main effects only)
    main_effect_mod  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
# One using head circumference, length, sex, and all interactions (including the three-way interaction) between these.
    interaction_mod     = map(train, \(df) lm(bwt ~ bhead + blength + babysex + 
                                        bhead*blength +
                                        bhead*babysex +
                                        blength*babysex +
                                        bhead * blength * babysex, data = df))
   ) |> 
  mutate(
    rmse_mine = map2_dbl(mine_model, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_main = map2_dbl(main_effect_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_interaction = map2_dbl(interaction_mod, test, \(mod, df) rmse(model = mod, data = df)))
```

```{r}
# compare the distribution of RMSE in plots
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin() + theme_bw() + 
  labs(x = "Model", y = "RMSE", 
       title = "Models vs. RMSE")
  
```

From the plot, we can see that our original model has the lowest RMSE, so we think that our model might be better compared to other two models in our senarios. 

## Problem 1 Solutions
```{r q1_data_cleaning}
homicide_df = 
  read_csv("DATA/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

